# Creating the layers {#building}

This chapter will cover the necessary steps to make layers which will be visualised in the app:

* kriging
* spatial joins
* aggregating interpolations within polygons


## Kriging

Kriging is an interpolation method that we use for iTRAQI. We pass observed values with known outcomes and coordinates and use kriging to get predicted values for new coordinates (the rest of Queensland).


### Data
First, we will download the data that we used for acute care travel time. Each row in the data has a coordinate (x,y) and outcome that we will be using for interpolation (time)

Table \@ref(tab:data-for-kriging) and figure \@ref(fig:map-of-data-for-kriging) show a preview of the data that we will be using.

```{r data-for-kriging, message=FALSE, warning=FALSE}
library(tidyverse)
library(leaflet)

save_dir <- "input"
githubURL <- glue::glue("https://raw.githubusercontent.com/RWParsons/iTRAQI_app/main/input/QLD_locations_with_RSQ_times_20220518.csv")
download.file(githubURL, file.path(save_dir, "df_acute.csv"), method="curl")

df_acute <- read.csv(file.path(save_dir, "df_acute.csv")) %>%
  select(location, x, y, centre=acute_care_centre, time=acute_time)

knitr::kable(
  head(df_acute, 10), caption = 'A preview of the data used for kriging',
  booktabs = TRUE
)
```

```{r map-of-data-for-kriging, fig.cap='leaflet map with locations', out.width='100%', fig.width=12, warning=FALSE}
leaflet() %>%
  addProviderTiles("CartoDB.VoyagerNoLabels") %>%
  addCircleMarkers(
    lng=df_acute$x, 
    lat=df_acute$y,
    popup=glue::glue( # customise your popups with html tags
      "<b>Location: </b>{df_acute$location}<br>",
      "<b>Time to acute care (minutes): </b>{df_acute$time}"),
    radius=2, fillOpacity=0,
  )
```

We will convert our data.frame into a spatial data.frame and load the gstat package as we will be using it for the kriging (`gstat::krige()`).

```{r}
library(sp)
library(gstat)
library(sf)

coordinates(df_acute) <- ~ x + y
```

### Making a grid of values for interpolation

Another key ingredient to do kriging is to have a grid of coordinates for which we want predictions (QLD).
The code below achieves this by creating a grid across all coordinates of QLD and keeping only those which intersect with the QLD boundary polygon. The initial grid contains coordinates for all combinations of latitudes and longitudes in QLD (which includes a lot of water of the north east for which we don't need interpolated values). Figure \@ref(fig:map-of-coord-grid) shows the initial grid made using `sp::makegrid()` in blue and the intersect between this and the QLD boundary in orange.  We will use the values which are within the QLD boundary for kriging. 

The cellsize we use here is large to save computation time (and to highlight a problem that we will come across very soon). This controls the resolution of the interpolation - the smaller the cellsize, the greater the spatial resolution. This is in degrees units (0.1 degree = 11.1km) so only having one prediction for every 11.1km² in QLD may mean that we miss out on some valuable information!

```{r map-of-coord-grid, fig.cap='coordinates that we will use for kriging (initial grid in blue and those than intersect with QLD boundary in orange)', out.width='100%', fig.width=12, warning=FALSE}
aus <- raster::getData('GADM', path="input", country = 'AUS', level = 1)
qld_boundary <- aus[aus$NAME_1 == "Queensland",]
qld_boundary_sf <- st_as_sfc(qld_boundary)

cellsize <- 0.05
grid <- makegrid(qld_boundary, cellsize = cellsize)
pnts_sf <- st_as_sf(grid, coords = c('x1', 'x2'), crs = st_crs(qld_boundary))

pnts_in_qld <- st_intersection(pnts_sf, qld_boundary_sf) %>% 
  st_coordinates() %>%
  as.data.frame()

ggplot() + 
  geom_point(data=grid, aes(x1, x2), col="blue") +
  geom_point(data=pnts_in_qld, aes(X, Y), col="orange") + 
  coord_equal() +
  labs(
    x="Longitude",
    y="Latitude"
  )

```

### Kriging (finally)
Now we are ready to do the kriging. `gstat::krige()` requires that the `newdata` be of class `Spatial`, `sf`, or `stars`. Here, I specify the coordinates using `sp::coordinates()`. It also requires that you specify the variogram model within - here we use a circular model `vgm("Cir")` but there may be better choices for other data.

Figure \@ref(fig:map-kriged-acute) shows the map with the interpolated values from kriging.

```{r map-kriged-acute, fig.cap='coordinates that we will use for kriging (initial grid in blue and those than intersect with QLD boundary in orange)', out.width='100%', fig.width=12}
lzn_vgm <- variogram(time ~ 1, df_acute)
lzn_fit <- fit.variogram(lzn_vgm, model=vgm("Sph"))

coordinates(pnts_in_qld) <- ~ X + Y

kriged_layer <-
  krige(
    formula=time ~ 1, 
    locations=df_acute,
    newdata=pnts_in_qld,
    model=lzn_fit
  ) %>%
  as.data.frame()

ggplot(data=kriged_layer, aes(X, Y, col=var1.pred)) + 
  geom_point() +
  scale_colour_gradientn(colors=c("yellow", "orange", "red", "black")) +
  coord_equal() +
  labs(
    x="Longitude",
    y="Latitude"
  )
```

### Making rasters

Now we can turn our grid of interpolated values into the rasters that we can then use in a leaflet map. We use the `raster` package. Figure \@ref(leaflet-map-raster) shows our kriged output as a raster on a leaflet map, the same type of objects as what's used in iTRAQI.

```{r leaflet-map-raster, fig.cap='coordinates that we will use for kriging (initial grid in blue and those than intersect with QLD boundary in orange)', out.width='100%', fig.width=12, warning=FALSE}
raster_layer <- raster::rasterFromXYZ(kriged_layer, crs=4326, res=0.05)
raster_layer <- raster::raster(raster_layer, layer=1) # layer=1 to select the prediction values rather than the variance

leaflet() %>%
  addProviderTiles("CartoDB.VoyagerNoLabels") %>%
  addRasterImage(x=raster_layer, colors="YlOrRd")
```

## Polygons and aggregations

We are going to download our polygons from the Australian Bureau of Statistics.

The link to the downloads page for the 2016 Australian Statistical Geography Standard (ASGS) files are [here](https://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/1270.0.55.001July%202016?OpenDocument) and the particular file that we are going to download is the 'Queensland Mesh Blocks ASGS Ed 2016 Digital Boundaries in ESRI Shapefile Format'. 
You will have to download the zipped file and unzip it somewhere locally. I've done so and saved it in the same directory as the other downloaded files and unzipped it into a folder there called 'qld_shape'. Having done that, I can import it using `st_read()`


```{r}
qld_SAs2016 <- st_read(file.path(save_dir, "qld_shape/MB_2016_QLD.shp"))

head(qld_SAs2016)
```

This data has polygons for every Statistical Area level 1 (SA1) in Queensland but also details the SA2, SA3, and SA4 that that area is within. If we want to only use SA1's then we are fine to use the data here, but if we want to use these higher levels too, then we would either need (1) make a new object with dissolved boundaries within that higher level or (2) download more files from the ABS for those specific levels and filter to keep only Queensland. These files that we could use, say for SA2's are called 'Statistical Area Level 2 (SA2) ASGS Ed 2016 Digital Boundaries in ESRI Shapefile Format', available at that same [link](https://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/1270.0.55.001July%202016?OpenDocument).

Since it's easy to filter, and reading this book is about learning new things (and my github repository is limited to 100mb), I'll show you the first approach that aggregates polygons within these higher levels.

Before we make a function to aggregate within different levels, I'm going to rename the columns in the object so that they're all named consistently - you may have noticed the unique identifier for SA1's is called 'SA1_MAIN16' whereas for SA3's it's called 'SA3_CODE16'. I prefer 'CODE'.
```{r}

qld_SAs2016 <-
  rename(qld_SAs2016, SA1_CODE16=SA1_MAIN16, SA2_CODE16=SA2_MAIN16)

```


### Dissolving polygons to get SA2s

The function below will dissolve the boundaries for all the polygons within the SA-level that we pick. The work here is done by `rmapshaper::ms_dissolve()`. I'll use this to make separate objects for SA2s and SA3s. Since this returns back only the geometry of the polygon and the name, I'll make the same change for my SA1s. By selecting only the code, I get the object with the code AND the geometry - unless I transform the object into a data.frame first, it will always keep the geometry.
```{r}

aggregate_by_SA <- function(qld_sf, SA_number){
  sa_main <- glue::glue('SA{SA_number}_CODE16')
  if(!sa_main %in% names(qld_sf)) return(message(sa_main, " was not found in polygon layer"))
  message(glue::glue('----- grouping polygons within SA{SA_number} -----'))
  rmapshaper::ms_dissolve(qld_sf, sa_main)
}


qld_SA2s <- aggregate_by_SA(qld_sf=qld_SAs2016, SA_number=2)
qld_SA3s <- aggregate_by_SA(qld_sf=qld_SAs2016, SA_number=3)

qld_SA1s <- qld_SAs2016 %>% select(SA1_CODE16)
head(qld_SA1s)
head(qld_SA2s)

```

There are some empty geometries here, so we find (and then remove) these using `st_is_empty()`.
```{r}
qld_SA1s <- qld_SA1s[!st_is_empty(qld_SA1s), , drop=FALSE]
qld_SA2s <- qld_SA2s[!st_is_empty(qld_SA2s), , drop=FALSE]
qld_SA3s <- qld_SA3s[!st_is_empty(qld_SA3s), , drop=FALSE]
```

Run the code to become impatient and find out how long it takes leaflet to display such a detailed polygon layer.

```{r, eval=F}

leaflet() %>%
  addTiles() %>%
  addPolygons(
    data=qld_SA1s,
    fillColor="Orange",
    color="black",
    weight=1
  )
```


### Simplifying polygons to reduce rendering time with leaflet
We need to do something about this - fortunately, we don't need all the incredible amounts of detail in the polygons for our map, so we can simplify them using `rmapshaper::ms_simplify()`.
Simplifying the polygons can take a few minutes but it makes the maps much faster to display.

```{r}
qld_SA1s <- rmapshaper::ms_simplify(qld_SA1s, keep=0.03)
qld_SA2s <- rmapshaper::ms_simplify(qld_SA2s, keep=0.03)
qld_SA3s <- rmapshaper::ms_simplify(qld_SA3s, keep=0.03)

leaflet() %>%
  addTiles() %>%
  addPolygons(
    data=qld_SA1s,
    fillColor="yellow",
    color="black",
    weight=1,
    group="SA1"
  ) %>%
  addPolygons(
    data=qld_SA2s,
    fillColor="orange",
    color="black",
    weight=1,
    group="SA2"
  ) %>%
  addPolygons(
    data=qld_SA3s,
    fillColor="red",
    color="black",
    weight=1,
    group="SA3"
  ) %>%
  addLayersControl(
    position="topright",
    baseGroups=c("SA1", "SA2", "SA3"),
    options=layersControlOptions(collapsed = FALSE)
  )

```


### Spatial joins and aggregations




